---
title: "Assignment Week4"
author: "Himanshu Singh"
date: "August 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Dataset 
In this section, we download the data files from the Internet and load them into two data frames. We ended up with a training dataset and a 20 observations testing dataset that will be submitted to Coursera.

```{r echo=FALSE}

library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
dt_training <- read.csv("./pml-training.csv", na.strings=c("NA","#DIV/0!",""))
dt_testing <- read.csv("./pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```   

## Cleaning the Data 
In this section, we will remove all columns that contains NA and remove features that are not in the testing dataset. The features containing NA are the variance, mean and standard devition (SD) within each window for each feature. Since the testing dataset has no time-dependence, these values are useless and can be disregarded. We will also remove the first 7 features since they are related to the time-series or are not numeric.

```{r echo=FALSE}
features <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]

# Only use features used in testing cases.
dt_training <- dt_training[,c(features,"classe")]
dt_testing <- dt_testing[,c(features,"problem_id")]

dim(dt_training)
dim(dt_testing)
```   

## Partitioning the Dataset 
Following the recommendation in the course Practical Machine Learning, we will split our data into a training data set (60% of the total cases) and a testing data set (40% of the total cases; the latter should not be confused with the data in the pml-testing.csv file). This will allow us to estimate the out of sample error of our predictor.

```{r echo=FALSE}
set.seed(12345)

inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]

dim(training)
dim(testing)
```   

## Building the Decision Tree Model 
Using Decision Tree, we shouldn't expect the accuracy to be high. In fact, anything around 80% would be acceptable.

```{r echo=FALSE}
modFitDT <- rpart(classe ~ ., data = training, method="class")
fancyRpartPlot(modFitDT)
```   

### Predicting with the Decision Tree Model
```{r echo=FALSE}
set.seed(12345)
prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```   

## Building the Random Forest Model
Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.
```{r echo=FALSE}
set.seed(12345)
modFitRF <- randomForest(classe ~ ., data = training, ntree = 1000)
```   
### Predicting with the Random Forest Model

```{r echo=FALSE}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```
## Predicting on the Testing Data (pml-testing.csv)
### Decision Tree Prediction

```{r echo=FALSE}
predictionDT <- predict(modFitDT, dt_testing, type = "class")
predictionDT
```   

### Random Forest Prediction

```{r echo=FALSE}
predictionRF <- predict(modFitRF, dt_testing, type = "class")
predictionRF
```   

## Submission file
As can be seen from the confusion matrix the Random Forest model is very accurate, about 99%. Because of that we could expect nearly all of the submitted test cases to be correct. It turned out they were all correct.

Prepare the submission.

```{r echo=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionRF)
```   




